Formal verification has advanced to the point that developers can verify the correctness of small, critical modules.  Unfortunately, despite considerable efforts, determining if a ``verification'' verifies what the author intends is still difficult.  Previous approaches are difficult to understand and often limited in applicability.  Developers need verification coverage in terms of the software they are verifying, not model checking diagnostics.  We propose a methodology to allow developers to determine (and correct) what it is that they have verified, and tools to support that methodology.  Our basic approach is based on a novel variation of mutation analysis and the idea of verification driven by falsification.  We use the CBMC model checker to show that this approach is applicable not only to simple data structures and sorting routines, and verification of a routine in Mozilla's JavaScript engine, but to understanding an ongoing effort to verify the Linux kernel Read-Copy-Update (RCU) mechanism.  Moreover, we show that despite the probabilistic nature of random testing and the tendency to incompleteness of testing as opposed to verification, the same techniques, with suitable modifications, apply to automated test generation as well as to formal verification.  In essence, it is the  number of surviving mutants that drives the scalability of our methods, not the underlying method for detecting faults in a program.  From the point of view of a Popperian analysis where an unkilled mutant is a weakness (in terms of its falsifiability) in a ``scientific theory'' of program behavior, it is only the number of weaknesses to be examined by a user that is important.